---
title: "RNA-Seq Manuscript Supplementary Material"
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: rmd-latex-ms.tex
# author:
# - name: Peng Zhou
#  affiliation: University of Minnesota
# - name: Nathan Springer
#  affiliation: University of Minnesota
# abstract: "This document provides an introduction to R Markdown, argues for its..."
# keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
# bibliography: ~/Dropbox/master.bib
# biblio-style: apsr
header-includes:
- \usepackage{caption}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---
\pagenumbering{gobble}

\pagebreak
```{r echo = FALSE}
require(tidyverse)
require(knitr)
require(kableExtra)
options(knitr.kable.NA = '')
dirw = '/home/springer/zhoux379/projects/briggs/data/41.qc' 
ft = file.path(dirw, "00.table.RData")
load(ft)
kable(t_sam, format = 'latex', longtable = T, booktabs = T,
  format.args = list(big.mark = ","), 
  caption = "Table S1. Samples used in this study.") %>%
  kable_styling(latex_options = c("repeat_header", "hold_position"), font_size=8)
```

\pagebreak
```{r echo = FALSE}
options(knitr.kable.NA = 0)
dirw = '/home/springer/zhoux379/projects/briggs/data/49.coop' 
fx = file.path(dirw, "51.go.enrich.RData")
load(fx)
kable(tx, format = 'latex', booktabs = T,
  caption = "Table S2. Enriched Gene Ontology (GO) terms for identified gene sets.") %>%
  kable_styling("striped", "hold_position", full_width = F) %>%
  column_spec(1, bold=T) %>%
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "top", row_group_label_position = 'stack')
```


\pagebreak
![Figure S1]($briggs/data/41.qc/01.readmapping.pdf){width=90%}

\pagebreak
![Figure S2]($briggs/data/41.qc/07.cpm.cor.pdf){width=100%}

\pagebreak
![Figure S3]($briggs/data/41.qc/09.fpkm.pca.pdf){width=100%}

\pagebreak
![Figure S4]($briggs/data/41.qc/12.cpm.den.pdf){width=100%}

\pagebreak
![Figure S5]($briggs/data/49.coop/25.hDE.pdf){width=100%}

\pagebreak
![Figure S6]($briggs/data/49.coop/25.Dom.sup.pdf){width=100%}

\pagebreak
![Figure S7]($briggs/data/42.de/15.density.doa.pdf){width=100%}

\pagebreak
![Figure S8]($briggs/data/49.coop/25.Reg.sup.pdf){width=100%}

\pagebreak
![Figure S9]($briggs/data/49.coop/25.Reg2.pdf){width=70%}

\pagebreak
![Figure S10]($briggs/data/49.coop/85.eQTL.pdf){width=100%}
